A meta-analysis of mental rotation ability in the first years of life

### 2022-05-05 code/analysis.Rmd
Update regression results

In the present study, we meta-analyzed `r descs$n_experiments` effect sizes derived from looking times in mental rotation tasks conducted by `r descs$n_infants` infants (`r print_perc(descs$percent_female)` female) aged `r descs$min_age_months` to `r descs$max_age_months` months (mean age of `r descs$mean_age_weighted`; Supplementary Table 1). All tasks were embedded either into habituation experiments (`r descs$n_experiments_habituation`) or violation of expectation experiments (`r descs$n_experiments_voe`) [@fantz1964; @baillargeon1985]. These experiments comprised real world stimuli (e.g., toy objects [@antrilli2016]), three-dimensional digital stimuli (e.g., cube figures [@moore2008]), or two-dimensional digital stimuli (e.g., digits [@quinn2008]). In habituation experiments, infants repeatedly saw an object until their looking times declined before they were presented with a mirror image of the object. Longer looking times at the mirror image were taken as evidence that an infant still recognized the familiar object after rotation through the new orientation. In violation of expectation experiments, infants were habituated to an object that was revolving repeatedly through a certain angle. Then, either the familiar object or an unseen object was shown while they were revolving through a certain angle. Subsequently, the object disappeared behind an occluder. Finally, the occluder was removed and either the object or its mirror object were shown. Larger differences in looking times for the familiar object versus the unseen mirror image (at the new angle) were taken as evidence that an infant still recognized the familiar object after rotation through the new angle. In contrast, looking times for both objects were expected to be similar if an infant did not recognize the familiar object from the new angle.

@@

# Extract posterior draws for the meta-analytic effects
spread_draws(
  res_brm,
  b_Intercept, sd_article__Intercept, `sd_article:experiment__Intercept`,
  ndraws = NULL
) %>%
  # Compute variances and ICC
  mutate(
    intercept = b_Intercept,
    sigma_article = sd_article__Intercept,
    sigma_experiment = `sd_article:experiment__Intercept`,
    sigma2_article = sigma_article^2,
    sigma2_experiment = sigma_experiment^2,
    sigma2_total = sigma2_article + sigma2_experiment,
    icc = sigma2_article / sigma2_total,
    .keep = "unused"
  ) -> draws_brm

@@

(ref:fig1-caption) Mental rotation performance. A Bayesian three-level meta-analysis provided evidence for mental rotation ability in infants. White squares depict the effect sizes (Hedges’ g) for infants’ mental rotation performance in all individual experiments and black lines depict their 95% confidence intervals. Gray bars indicate the 95% and 50% Bayesian credible intervals (CrI). These intervals got shrunk towards the meta-analytic effect size because of partial pooling which regularized the impact of experiments with small sample sizes or unrealistically large effect sizes to prevent overfitting the model. The last line shows the meta-analytic effect size (black dot) together with its 95% CrI (black line) and its posterior distribution (gray curve).

@@

@@

# Compute by-experiment variance
var_experiment_reg <- as.matrix(
  res_brm_reg,
  variable = "sd_article:experiment__Intercept"
)^2
mean_qi(var_experiment_reg)

# Compute by-article variance
var_article_reg <- as.matrix(
  res_brm_reg,
  variable = "sd_article__Intercept"
)^2
mean_qi(var_article_reg)

# # Compute intra-class correlation
# var_total_reg <- var_experiment_reg + var_article_reg
# icc_reg <- var_article_reg / var_total_reg
# mean_qi(icc_reg)

# # Compute reduction in heterogeneity compared to the original meta-analysis
# 1 - mean(var_total_reg / var_total)
```

As a next step, we conducted a meta-regression analysis to test if the gender of the infants, their age, or the type of mental rotation task was related to mental rotation performance (Fig. 2).
Indeed, experiments with all-male samples revealed larger looking time differences than experiments with all-female samples, b = 0.48, 95% CrI [0.06, 0.91].
The probability for this effect being larger than zero was 98.5%.
We found no difference between mixed-gender and all-female samples, b = 0.13, 95% CrI [-0.20, 0.52] (75.4% probability of an effect larger than zero).
Additionally, mean age was not related to mental rotation performance, with a change per year of b = 0.05, 95% [-0.58, 0.68].
We also did not detect an interaction between gender and age ([females - mixed] × age: b = 0.02, 95% CrI [-0.81, 0.84]; [males - females] × age: b = 0.06, 95% CrI [-0.84, 0.96]).
Finally, there was weak evidence that violation of expectation tasks yielded larger effects than habituation tasks, b = 0.26, 95% [-0.08, 0.58].
The probability for this effect being greater than zero was 93.5%.

@@

(ref:fig2-caption) Effects of gender, age, and task. a Squares show the effect size (Hedges' g) for infants' mental rotation performance in each of the 59 individual experiments. Squares are color-coded according to the type of habituation task and the gender of the infants (blue = habituation task, all-female sample, yellow = habituation task, all-male sample, red = habituation task, mixed-gender sample, purple = violation of expectation [VoE] task, mixed-gender sample). Lines indicate the best-fit regression estimates according to a Bayesian three-level meta-regression model and gray ribbons indicate their corresponding 95% and 50% credible interval (CrI). b Fixed effect estimates obtained from the Bayesian three-level meta-regression model are depicted as black dots together with their 95% Crl (thin black lines) and 50% Crl (thick black lines). Gray curves indicate the posterior distribution for each effect.

>>>>

In the present study, we meta-analyzed `r descs$n_experiments` effect sizes derived from looking times in mental rotation tasks conducted by `r formatC(descs$n_infants, big.mark = ",")` infants (`r print_perc(descs$percent_female)` female) aged `r descs$min_age_months` to `r descs$max_age_months` months (mean age of `r descs$mean_age_weighted`; Supplementary Table 1). All tasks were embedded either into habituation experiments (`r descs$n_experiments_habituation`) or violation of expectation experiments (`r descs$n_experiments_voe`) [@fantz1964; @baillargeon1985]. These experiments comprised real world stimuli (e.g., toy objects [@antrilli2016]), three-dimensional digital stimuli (e.g., cube figures [@moore2008]), or two-dimensional digital stimuli (e.g., digits [@quinn2008]). In habituation experiments, infants repeatedly saw an object until their looking times declined before they were presented with a mirror image of the object. Longer looking times at the mirror image were taken as evidence that an infant still recognized the familiar object after rotation through the new orientation. In violation of expectation experiments, infants were habituated to an object that was revolving repeatedly through a certain angle. Then, either the familiar object or an unseen object was shown while they were revolving through a certain angle. Subsequently, the object disappeared behind an occluder. Finally, the occluder was removed and either the object or its mirror object were shown. Larger differences in looking times for the familiar object versus the unseen mirror image (at the new angle) were taken as evidence that an infant still recognized the familiar object after rotation through the new angle. In contrast, looking times for both objects were expected to be similar if an infant did not recognize the familiar object from the new angle.

@@

# Extract posterior draws for the meta-analytic effects
spread_draws(res_brm, `b_.*`, `sd_.*`, regex = TRUE, ndraws = NULL) %>%
  # Compute variances and ICC
  mutate(
    intercept = b_Intercept,
    sigma_article = sd_article__Intercept,
    sigma_experiment = `sd_article:experiment__Intercept`,
    sigma2_article = sigma_article^2,
    sigma2_experiment = sigma_experiment^2,
    sigma2_total = sigma2_article + sigma2_experiment,
    icc = sigma2_article / sigma2_total,
    .keep = "unused"
  ) -> draws_brm

@@

(ref:fig1-caption) **Mental rotation performance.** A Bayesian three-level meta-analysis provided evidence for mental rotation ability in infants. White squares depict the effect sizes (Hedges’ $g$) for infants’ mental rotation performance in all individual experiments and black lines depict their 95% confidence intervals. Gray bars indicate the 95% and 50% Bayesian credible intervals (CrI). These intervals got shrunk towards the meta-analytic effect size because of partial pooling which regularized the impact of experiments with small sample sizes or unrealistically large effect sizes to prevent overfitting the model. The last line shows the meta-analytic effect size (black dot) together with its 95% CrI (black line) and its posterior distribution (gray curve).

@@

# Get posterior probability mass > 0 for all regression weights
(hyps_brm_reg <- list(
  intercept = hypothesis(res_brm_reg, "Intercept > 0")$hypothesis,
  female_mixed = hypothesis(res_brm_reg, "gender2M1 < 0")$hypothesis,
  male_female = hypothesis(res_brm_reg, "gender3M2 > 0")$hypothesis,
  age = hypothesis(res_brm_reg, "age_c > 0")$hypothesis,
  voe_habituation = hypothesis(res_brm_reg, "task2M1 > 0")$hypothesis,
  female_mixed_age = hypothesis(res_brm_reg, "gender2M1:age_c > 0")$hypothesis,
  male_female_age = hypothesis(res_brm_reg, "gender3M2:age_c > 0")$hypothesis
))

@@

# Extract posterior draws for the meta-analytic effects
spread_draws(res_brm_reg, `b_.*`, `sd_.*`, regex = TRUE, ndraws = NULL) %>%
  # Compute variances and ICC
  mutate(
    intercept = b_Intercept,
    female_mixed = b_gender2M1,
    male_female = b_gender3M2,
    age = b_age_c,
    voe_habituation = b_task2M1,
    female_mixed_age = `b_gender2M1:age_c`,
    male_female_age = `b_gender3M2:age_c`,
    sigma_article = sd_article__Intercept,
    sigma_experiment = `sd_article:experiment__Intercept`,
    sigma2_article = sd_article__Intercept^2,
    sigma2_experiment = `sd_article:experiment__Intercept`^2,
    sigma2_total = sigma2_article + sigma2_experiment,
    icc = sigma2_article / sigma2_total,
  ) -> draws_brm_reg

# Summarize as mean and 95% credible intervals
(summ_brm_reg <- as.list(mean_qi(draws_brm_reg)))
```

As a next step, we conducted a meta-regression analysis to test if the gender of the infants, their age, or the type of mental rotation task was related to mental rotation performance (Fig. 2).
Indeed, experiments with all-male samples revealed larger looking time differences than experiments with all-female samples, $b = `r summ_brm_reg$male_female`$, 95% CrI [`r summ_brm_reg$male_female.lower`, `r summ_brm_reg$male_female.upper`].
The probability for this effect being larger than zero was `r print_perc(hyps_brm_reg$male_female$Post.Prob)`.
We found no difference between mixed-gender and all-female samples, $b = `r -summ_brm_reg$female_mixed`$, 95% CrI [`r -summ_brm_reg$female_mixed.lower`, `r -summ_brm_reg$female_mixed.upper`] (`r print_perc(hyps_brm_reg$female_mixed$Post.Prob)` probability of an effect larger than zero).
Additionally, mean age was not related to mental rotation performance, with a change per year of $b = `r summ_brm_reg$age`$, 95% [`r summ_brm_reg$age.lower`, `r summ_brm_reg$age.upper`].
We also did not detect an interaction between gender and age ([females - mixed] × age: $b = `r summ_brm_reg$female_mixed_age`$, 95% CrI [`r summ_brm_reg$female_mixed_age.lower`, `r summ_brm_reg$female_mixed_age.upper`]; [males - females] × age: $b = `r summ_brm_reg$male_female_age`$, 95% CrI [`r summ_brm_reg$male_female_age.lower`, `r summ_brm_reg$male_female_age.upper`]).
Finally, there was weak evidence that violation of expectation tasks yielded larger effects than habituation tasks, $b = `r summ_brm_reg$voe_habituation`$, 95% [`r summ_brm_reg$voe_habituation.lower`, `r summ_brm_reg$voe_habituation.upper`].
The probability for this effect being greater than zero was `r print_perc(hyps_brm_reg$voe_habituation$Post.Prob)`.

@@

(ref:fig2-caption) **Effects of gender, age, and task.** **a,** Squares show the effect size (Hedges’ g) for infants’ mental rotation performance in each of the 59 individual experiments. Squares are color-coded according to the type of habituation task and the gender of the infants (blue = habituation task, all-female sample, yellow = habituation task, all-male sample, red = habituation task, mixed-gender sample, purple = violation of expectation [VoE] task, mixed-gender sample). Lines indicate the best-fit regression estimates according to a Bayesian three-level meta-regression model and gray ribbons indicate their corresponding 95% and 50% credible interval (CrI). **b,** Fixed effect estimates obtained from the Bayesian three-level meta-regression model are depicted as black dots together with their 95% Crl (thin black lines) and 50% Crl (thick black lines). Gray curves indicate the posterior distribution for each effect.
